Extended Data

### 2022-03-12 doc/manuscript/Dow_manuscript_Nature.Rmd
minor

<!--summary, setting expectation of extra C-->
Together, our results demonstrate that warmer spring temperatures in the temperate deciduous forests of eastern North America advance the timing of stem growth but have little effect on total annual growth (Figs. `r schematic_summary`-`r tree_rings`). 
The observed advance in stem growth under warmer springs parallels phenological advances observed for canopy foliage (Fig. `r DOY_timing`a-b)[@keenan_net_2014;@finzi_carbon_2020;@stuble_plant_2021] and NEE[@keenan_net_2014;@finzi_carbon_2020]. 
However, inconsistent with the concept that an earlier start to growth would increase $ANPP_{woody}$, we demonstrate that warmer springs either hasten the deceleration of stem expansion or otherwise fail to translate extended growing seasons into biologically significant increases in stem growth (Fig. `r schematic_summary`), and thereby have negligible effect on total annual growth for most species and locations (Fig. `r tree_rings`).
Our observations suggest that the deceleration of stem expansion, which occurs mid-summer near the time of peak canopy greenness (`r EDF_growth_curves_all`)[@finzi_carbon_2020;@etzold_number_2022], is likely driven by cues other than photosynthate limitation, such as water stress, daylength, or sink limitation[@zweifel_why_2021;@etzold_number_2022;@zani_increased_2020;@keenan_timing_2015;<!--@fu_variation_2014;-->@delpierre_wood_2016;@delpierre_temperate_2016].
Combined with observations that warming springs tend to lengthen the season of C sequestration[@refs], our observations imply a lengthening of the period from peak stem growth to the cessation of CO~2~ uptake by the ecosystem.
What happens with the extra C fixed during this period?

<!--extra C is probably not going to ANPP_woody-->
Additional C fixed in years with warm springs does not substantially augment stem diameter growth in the current or following year (`r EDF_tree_rings_TMAX_extended`), and is unlikely to augment $ANPP_{woody}$, which, although routinely calculated based on stem growth, can be partially decoupled from it through differences in wood density or C content[@delpierre_wood_2016].
Thus, it is theoretically possible that extra C is allocated to enhanced cell wall thickening, a process that lags behind stem expansion[@cuny_woody_2015], or to a higher ratio of high-density latewood to lower-density earlywood, resulting in formation of more C-dense wood in years with warm springs.
However, existing evidence indicates that warm springs have a neutral or negative effect on latewood width[<!--@tardif_influence_2006;-->@roibu_climatic_2020;@kern_multiple_2013], which is more strongly controlled by summer drought stress[<!--@tardif_influence_2006;-->@roibu_climatic_2020;@kern_multiple_2013], suggesting that a positive effect of warm springs on the total C content of annual rings is unlikely. 
Rather, the most likely explanation of the observed lack of correlation between spring temperatures and stem growth is that growth is sink-limited, such that increased GPP does not enhance $ANPP_{woody}$[@delpierre_wood_2016].

>>>>

<!--summary, setting expectation of extra C-->
Together, our results demonstrate that warmer spring temperatures in the temperate deciduous forests of eastern North America advance the timing of stem growth but have little effect on total annual growth (Figs. `r schematic_summary`-`r tree_rings`). 
The observed advance in stem growth under warmer springs parallels phenological advances observed for canopy foliage (Fig. `r DOY_timing`a-b)[@keenan_net_2014;@finzi_carbon_2020;@stuble_plant_2021] and NEE[@keenan_net_2014;@finzi_carbon_2020]. 
However, inconsistent with the concept that an earlier start to growth would increase $ANPP_{woody}$, we demonstrate that warmer springs either hasten the deceleration of stem expansion or otherwise fail to translate extended growing seasons into biologically significant increases in stem growth (Fig. `r schematic_summary`), and thereby have negligible effect on total annual growth for most species and locations (Fig. `r tree_rings`).
Our observations suggest that the deceleration of stem expansion, which occurs mid-summer near the time of peak canopy greenness (`r EDF_growth_curves_all`)[@finzi_carbon_2020;@etzold_number_2022], is likely driven by cues other than photosynthate limitation, such as water stress, daylength, or sink limitation[@zweifel_why_2021;@etzold_number_2022;@zani_increased_2020;@keenan_timing_2015;<!--@fu_variation_2014;-->@delpierre_wood_2016;@delpierre_temperate_2016].
Combined with observations that warming springs tend to lengthen the season of C sequestration[@lu_no_2022;@finzi_carbon_2020], our findings imply a lengthening of the period from peak stem growth to the cessation of CO~2~ uptake by the ecosystem.
What happens with the extra C fixed during this period?

<!--extra C is probably not going to ANPP_woody-->
Additional C fixation in years with warm springs[@keeling_increased_1996;@dragoni_evidence_2011;@keenan_net_2014;@crabbe_extreme_2016;@zhou_explaining_2016;@fu_maximum_2019;@finzi_carbon_2020] does not substantially augment stem diameter growth in the current or following year (`r EDF_tree_rings_TMAX_extended`), and is unlikely to augment $ANPP_{woody}$, which, although routinely calculated based on stem growth, can be partially decoupled from it through differences in wood density or C content[@delpierre_wood_2016].
It is theoretically possible that extra C is allocated to enhanced cell wall thickening, a process that lags behind stem expansion[@cuny_woody_2015], or to a higher ratio of high-density latewood to lower-density earlywood, resulting in formation of more C-dense wood in years with warm springs.
However, existing evidence indicates that warm springs have a neutral or negative effect on latewood width[<!--@tardif_influence_2006;-->@roibu_climatic_2020;@kern_multiple_2013], which is more strongly controlled by summer drought stress[<!--@tardif_influence_2006;-->@roibu_climatic_2020;@kern_multiple_2013], suggesting that a positive effect of warm springs on the total C content of annual rings is unlikely. 
Rather, the most likely explanation of the observed lack of correlation between spring temperatures and stem growth is that growth is sink-limited, such that increased GPP does not enhance $ANPP_{woody}$[@delpierre_wood_2016].
